---
title: "R Notebook"
output: html_notebook
---

```{r}
library(brms)
rstan_options(auto_write = TRUE)
options(mc.cores = 2)
```


get data
```{r}
load("pheno.geno.Rdata")
geno.center <- geno*2-1
```

test how to make large formula
```{r}
my.formula <- paste("high.RFR ~", paste(colnames(geno)[1:15],collapse="+"))
```

```{r}
brms1 <- brm(my.formula, data=pheno.geno,
            prior=c(
              set_prior("horseshoe(3,scale_global=0.1)",class="b"),
              set_prior("normal(3.5,1)",class="Intercept")
            ),
            control = list(adapt_delta = 0.8)
)
```


```{r}
scode1 <- make_stancode(my.formula, data=pheno.geno,
            prior=c(
              set_prior("horseshoe(3,scale_global=0.1)",class="b"),
              set_prior("normal(3.5,1)",class="Intercept")
            ),
            control = list(adapt_delta = 0.8)
)

scode1
```

```{r}
sdata1 <- make_standata(my.formula, data=pheno.geno,
            prior=c(
              set_prior("horseshoe(3,scale_global=0.1)",class="b"),
              set_prior("normal(3.5,1)",class="Intercept")
            )
)

names(sdata1)
str(sdata1)
```

```{r}
sdata_full <- list(
  N=as.integer(169),
  Y=pheno.geno$high.RFR,
  K=ncol(geno),
  X=geno,
  prior_only=as.integer(0)
)
```

45 seconds for 10, 450 for 100, 4500 for 1000, 9000 for 2000 = 2.5  hours?
```{r, eval=FALSE}
system.time(stan_full1 <- stan(
  model_code=scode1,
  data=sdata_full,
  verbose = TRUE,
# sample_file="stan_full1_samples",
#  diagnostic_file = "stan_full1_diagnostic_file",
  iter = 2000,
cores=2,
refresh=10,
))
```

Reading STAN boards, it seems that STAN doesn't do well with highly correlated predictors.  Maybe I should try doing PCA or MDS and then fit STAN on that...


```{r}
geno.pc <- prcomp(t(geno))
dim(geno.pc$rotation)
dim(geno.pc$x)
plot(geno.pc)
summary(geno.pc)
head(geno.pc$x)
head(geno.pc$rotation)
```
```{r}
pheno.pc <- as.data.frame(cbind(pheno,geno.pc$rotation))
head(pheno.pc[,1:20])
```

test how to make large formula
```{r}
my.formula2 <- paste("high.RFR ~", paste(grep("PC",colnames(pheno.pc),value=TRUE)[1:100],collapse="+"))
```

```{r}
brms2 <- brm(my.formula2, data=pheno.pc[1:120,],
            prior=c(
              set_prior("horseshoe(3,scale_global=0.5)",class="b"),
              set_prior("normal(3.5,1)",class="Intercept")
            ),
            iter=20000,
            warmup = 3000,
            control = list(adapt_delta = 0.9)
)
```
```{r}
shinystan::launch_shinystan(brms2,ask=FALSE)
summary(brms2)
```
```{r}
fitted <- as.data.frame(predict(brms2))
predicted <- as.data.frame(predict(brms2,pheno.pc[121:169,]))
plot(pheno.pc$high.RFR[1:120],fitted$Estimate)
cor(pheno.pc$high.RFR[1:120],fitted$Estimate)
plot(pheno.pc$high.RFR[121:169],predicted$Estimate)
cor(pheno.pc$high.RFR[121:169],predicted$Estimate)
```

```{r}
brms3 <- update(brms2,
            prior=c(
              set_prior("horseshoe(3,scale_global=1)",class="b"),
              set_prior("normal(3.5,1)",class="Intercept")
            ),
            iter=20000,
            warmup = 3000,
            control = list(adapt_delta = 0.9)
)
```

```{r}
summary(brms3)
```
```{r}
fitted <- as.data.frame(predict(brms3))
predicted <- as.data.frame(predict(brms3,pheno.pc[121:169,]))
plot(pheno.pc$high.RFR[1:120],fitted$Estimate)
cor(pheno.pc$high.RFR[1:120],fitted$Estimate)
plot(pheno.pc$high.RFR[121:169],predicted$Estimate)
cor(pheno.pc$high.RFR[121:169],predicted$Estimate)
```


```{r}
brms4 <- update(brms2,
            prior=c(
              set_prior("horseshoe(3,scale_global=.1)",class="b"),
              set_prior("normal(3.5,1)",class="Intercept")
            ),
            iter=20000,
            warmup = 3000,
            control = list(adapt_delta = 0.9)
)
```

```{r}
summary(brms4)
```

```{r}
fitted <- as.data.frame(predict(brms4))
predicted <- as.data.frame(predict(brms4,pheno.pc[121:169,]))
plot(pheno.pc$high.RFR[1:120],fitted$Estimate)
cor(pheno.pc$high.RFR[1:120],fitted$Estimate)
plot(pheno.pc$high.RFR[121:169],predicted$Estimate)
cor(pheno.pc$high.RFR[121:169],predicted$Estimate)
loo(brms2,brms3,brms4)
```
Ultimately all of these models are terrible at predicting the test set.  Probably because the individual important SNPs are not captured in the PCs